---
title: "Basic Estimation of Crop Yield Response (Econometrics)"
author:  "Jaeseok Hwang"
date: "Sep 2024"
format: html
bibliography: E00_econometrics_ref.bib
---

```{r}
#| include: false

library(data.table)
library(tidyverse)
library(mgcv)
library(parallel)
library(patchwork)
library(caret)
library(ggplot2)

```

## Estimation of Yield Respoonse to Nitrogen

### Example: Graph of Yield Response to Nitrogen Using M-B Function

The following R code generates sample data for nitrogen input, yield, and an error term. It then plots the observed data (green dots) along with the fitted M-B production function (red line).


```{r}

#| fig-height: 3
#| fig-width: 5
#| fig-align: center

# Generate sample data
set.seed(123)
nitrogen <- seq(0, 200, by = 5)  # Nitrogen input
Y_max <- 120  # Maximum yield
beta <- 0.03  # Nitrogen efficiency parameter
epsilon <- rnorm(length(nitrogen), mean = 0, sd = 5)  # Random error

# Calculate yield using M-B function
yield <- Y_max * (1 - exp(-beta * nitrogen)) + epsilon

# Create a data frame for plotting
data <- data.frame(Nitrogen = nitrogen, Yield = yield)

# Fit the M-B function
fit <- nls(Yield ~ Y_max * (1 - exp(-beta * Nitrogen)), 
           data = data, 
           start = list(Y_max = 120, beta = 0.03))

# Generate predicted values for plotting
data$Predicted_Yield <- predict(fit)

# Plot the observed data and fitted curve
ggplot(data, aes(x = Nitrogen, y = Yield)) +
  geom_point(color = "green", size = 2, alpha = 0.8) +  # Observed data
  geom_line(aes(y = Predicted_Yield), color = "red", size = 1)

```


### Interpretation of the Yield Response to Nitrogen Plot

The plot above demonstrates the relationship between nitrogen application and crop yield. The observed data (green dots) illustrate variability due to random error, while the red line represents the fitted Mitscherlich-Baule (M-B) function. This function effectively captures the diminishing returns to nitrogen and the asymptotic nature of yield, where yields plateau as nitrogen levels approach saturation. 

While the M-B production function is biologically intuitive, its parameters are nonlinear and require specialized estimation techniques, such as nonlinear least squares (as shown above). In practical applications, simpler functional forms, such as quadratic or quadratic plateau models, are often used for estimating yield response due to their ease of implementation within standard linear regression frameworks.

### Bridging the M-B Function and Linear Regression Models

To estimate yield response to nitrogen using regression models, a common approach is to employ a quadratic or quadratic plateau model. These models approximate the nonlinear relationship between nitrogen and yield while maintaining the simplicity of linear regression.

#### Quadratic Yield Response Model

The quadratic model assumes a concave relationship between nitrogen input and yield:

$$ Y = \beta_0 + \beta_1 N + \beta_2 N^2 + \epsilon $$

Where: <br> 
- $Y$ = Crop yield <br>
- $N$ = Nitrogen input <br>
- $\beta_0$, $\beta_1$, $\beta_2$ = Model parameters <br>
- $\epsilon$ = Error term

This model captures the diminishing returns to nitrogen through the quadratic term ($N^2$), with $\beta_2 < 0$. It has been widely used in empirical studies of nitrogen response due to its simplicity and flexibility.

Cite: @lory2003yield

#### Quadratic-Plateau Model

A variation of the quadratic model is the quadratic plateau model, which assumes that yields plateau after a certain level of nitrogen input:

$$ Y = \begin{cases} 
\beta_0 + \beta_1 N + \beta_2 N^2, & \text{if } N < N_{\text{critical}} \\
Y_{\text{max}}, & \text{if } N \geq N_{\text{critical}}
\end{cases} $$

Where: <br>
- $N_{\text{critical}}$ = Critical nitrogen rate at which the yield plateaus <br>
- $Y_{\text{max}}$ = Maximum attainable yield <br>

This model incorporates biological realism similar to the M-B function but remains computationally simpler for estimation.

Cite:  @cerrato1990comparison

#### Linear Regression Approach to Estimate Yield Response

To estimate yield response using regression, we start with the quadratic model as a foundation. The following R code demonstrates a simple implementation using synthetic data:

```{r}
#| fig-height: 3
#| fig-width: 5
#| fig-align: center

# Generate synthetic data
set.seed(123)
nitrogen <- seq(0, 200, by = 5)
beta_0 <- 40  # Intercept
beta_1 <- 0.8  # Linear nitrogen effect
beta_2 <- -0.002  # Quadratic nitrogen effect
epsilon <- rnorm(length(nitrogen), mean = 0, sd = 5)  # Random error

# Quadratic yield data
yield <- beta_0 + beta_1 * nitrogen + beta_2 * nitrogen^2 + epsilon
data <- data.frame(Nitrogen = nitrogen, Yield = yield)

# Fit a quadratic regression model
quad_fit <- lm(Yield ~ Nitrogen + I(Nitrogen^2), data = data)

# Plot observed data and fitted curve
ggplot(data, aes(x = Nitrogen, y = Yield)) +
  geom_point(color = "green", size = 2, alpha = 0.8) +  # Observed data
  stat_smooth(method = "lm", formula = y ~ poly(x, 2), se = FALSE, color = "red") +  # Fitted curve
  labs(
    title = "Quadratic Yield Response to Nitrogen",
    x = "Nitrogen Input (kg/ha)",
    y = "Yield (kg/ha)"
  ) +
  theme_minimal()
```