---
title: "Yield Distribution"
format:
  html:
    toc: true
---

```{r}
#| label: setup
#| include: false
knitr::opts_chunk$set(message = FALSE, warning = FALSE)
```

```{r}
# -- Configure repo + assets (EDIT these paths)
repo_user <- "jaeseokh"
repo_name <- "yield_distribution"
branch    <- "main"

figure_path <- "results/figures/main_figure.png"
table_path  <- "results/tables/summary.csv"

base_raw <- sprintf("https://raw.githubusercontent.com/%s/%s/%s/",
                    repo_user, repo_name, branch)
url_fig <- paste0(base_raw, figure_path)
url_tbl <- paste0(base_raw, table_path)
```

::: {.hero-mini}

# Why it matters

Means hide risk. Managers need the **whole distribution** to price decisions.
:::

## At a glance

* **Main problem**: Averages understate tail risk.
* **Main question**: How do N × weather shift moments/tails?
* **Finding / limitation**: Tails respond differently; calibration ongoing.
* **Long‑term solution & goal**: Central/raw‑moment regressions → MaxEnt densities.

## Key Result

### Main Figure

```{r}
#| fig-cap: "Latest main figure from the repo"
knitr::include_graphics(url_fig)
```

**Takeaway:** Two sentences about risk‑sensitive insights from the figure.

### Main Table (top lines)

```{r}
library(readr); library(dplyr)
try({
  tbl <- read_csv(url_tbl, show_col_types = FALSE)
  head(tbl, 8)
}, silent = TRUE)
```

**Takeaway:** Two sentences on moment shifts and implications.

## Workflow Roadmap

<div class="stepper">
  <div class="step">Moment models</div>
  <div class="step inprog">MaxEnt</div>
  <div class="step">Backtests</div>
  <div class="step">Docs</div>
</div>

## Tech & Data

**Tech**: R, MaxEnt, NB/GLM, CAR

**Data**: OFPE + weather/soil features.

<div class="actions">
  <a class="btn primary" href="https://github.com/jaeseokh/yield_distribution">GitHub Repository</a>
</div>
```
