---
title: "Multi‑Field OFPE Value"
format:
  html:
    toc: true
---

```{r}
#| label: setup
#| include: false
knitr::opts_chunk$set(message = FALSE, warning = FALSE)
```

```{r}
# -- Configure repo + assets (EDIT these paths to your actual files)
repo_user <- "jaeseokh"
repo_name <- "value_multifield_ofpe"
branch    <- "main"    # or the branch where results live

# Example: results/figures/main_figure.png and results/tables/summary.csv
figure_path <- "results/figures/main_figure.png"
table_path  <- "results/tables/summary.csv"

base_raw <- sprintf("https://raw.githubusercontent.com/%s/%s/%s/",
                    repo_user, repo_name, branch)
url_fig <- paste0(base_raw, figure_path)
url_tbl <- paste0(base_raw, table_path)
```

::: {.hero-mini}

# Why it matters

Pooling on‑farm precision experiments across fields can turn small local trials into **field‑scale, robust recommendations**.
:::

## At a glance

* **Main problem (why)**: Farmers run OFPEs, but gains are often siloed by field; pooling is hard.
* **Main question (focus)**: How to quantify cross‑field value while respecting heterogeneity?
* **Finding / limitation**: Early evidence of decision value; pooling sensitive to outliers.
* **Long‑term solution & goal**: Hierarchical pooling + diagnostics → robust field prescriptions.

## Key Result

### Main Figure

```{r}
#| fig-cap: "Latest main figure from the repo"
knitr::include_graphics(url_fig)
```

**Takeaway (2 sentences):**

* The pooled model identifies consistent response patterns across fields.
* Remaining variance suggests targeted pooling rules and outlier handling are pivotal.

### Main Table (top lines)

```{r}
library(readr); library(dplyr)
try({
  tbl <- read_csv(url_tbl, show_col_types = FALSE)
  head(tbl, 8)
}, silent = TRUE)
```

**Takeaway (2 sentences):**

* The summary highlights between‑field variability that affects expected gains.
* We prioritize shrinkage strategies where uncertainty is largest.

## Workflow Roadmap

<div class="stepper">
  <div class="step done">Data audit</div>
  <div class="step inprog">Pooling models</div>
  <div class="step">Diagnostics</div>
  <div class="step">Prescription</div>
  <div class="step">Docs</div>
</div>

## Tech & Data

**Tech**: R, Quarto, GLM/GLMM, Spatial/CAR, Bayes

**Data**: OFPE field trials (multi‑year, multi‑field).

<div class="actions">
  <a class="btn primary" href="https://github.com/jaeseokh/value_multifield_ofpe">GitHub Repository</a>
</div>
```
